\section{Introduction}

Autoregressive decoding in large language models (LLMs) is fundamentally
sequential: each output token requires a full forward pass through the
target model, capping single-stream throughput at the per-token kernel
latency. For 7\,B-class models served on a single consumer GPU, this is the
dominant cost in latency-sensitive applications such as code completion,
on-device assistants, and interactive chat. \emph{Speculative decoding}
addresses this bottleneck by interposing a small \emph{draft} model that
proposes $k$ candidate tokens, which the target then verifies in parallel
with a single forward pass; rejected tokens are re-sampled from the
target's residual distribution to preserve the output
distribution~\cite{Leviathan2023,Chen2023specsampling}.

The headline numbers for speculative decoding ($S\!\in\![2\times,5\times]$)
are typically reported on data-center GPUs (A100, H100, H200) using
unquantised drafters. On consumer hardware with quantised models the
picture is murkier: per-block bookkeeping overhead, dequantisation cost,
and asymmetric draft/target throughput can erase the algorithmic gain
entirely. Practitioners deploying speculative decoding on a single RTX
4090- or 5090-class GPU therefore lack quantitative guidance on which
$(\text{draft size}, k, \text{regime})$ to pick, and at what point the
linear-in-$k$ drafting cost itself becomes the bottleneck.

\subsection{Research questions}
\label{sec:rqs}

We address three research questions:
\begin{enumerate}\itemsep0pt
  \item[\textbf{RQ1.}] How much end-to-end speedup does vanilla
        speculative decoding deliver against autoregressive decoding on a
        single RTX~5090 with an \texttt{int8} 7\,B target?
  \item[\textbf{RQ2.}] How do draft model size and draft length jointly
        determine token acceptance and net latency?
  \item[\textbf{RQ3.}] How sensitive is the speed--quality trade-off to
        the sampling regime (deterministic vs.\ stochastic) and to task
        entropy?
\end{enumerate}

\subsection{Contributions}
\label{sec:contributions}

This paper makes the following contributions:
\begin{enumerate}\itemsep0pt
  \item A reproducible empirical protocol with frozen sample manifests on
        three benchmarks (GSM8K, MMLU subsets, CNN/DailyMail) and a
        14-configuration grid (Qwen2.5-7B target; 0.5B/1.5B drafts;
        $k\!\in\!\{4,8,16\}$; deterministic and stochastic regimes).
  \item Wall-clock and acceptance results showing a best speedup of
        $S{=}1.65\times$ at $k{=}8$ deterministic, and an empirical
        $\bar{B}_{\text{eff}}$ that tracks the geometric upper bound
        within $3\%$, identifying the linear-in-$k$ draft cost as the
        binding term once per-block overhead is amortised.
  \item \emph{DriftDiffuse}: a $\sim$30\,M-parameter bidirectional
        masked-diffusion drafter with a position-conditional drift
        schedule and MaskGIT-style iterative
        unmasking~\cite{Chang2022MaskGIT}, supporting both block-level
        and Leviathan-style token-level acceptance and producing $k$
        draft tokens in $n_{\text{denoise}}{\ll}k$ parallel steps.
  \item A break-even analysis that explains why parallel drafters become
        attractive at $k{\geq}8$ on consumer hardware, together with the
        complete training and evaluation pipeline distilled from the
        target's own outputs.
\end{enumerate}

The rest of the paper is organised as follows. Section~\ref{sec:related}
reviews related work. Section~\ref{sec:protocol} defines the experimental
protocol; Section~\ref{sec:metrics} the metrics. Section~\ref{sec:impl}
covers implementation. Section~\ref{sec:results} presents the autoregressive
results, and Section~\ref{sec:driftdiffuse} introduces DriftDiffuse.
Section~\ref{sec:discussion} discusses systems implications, and
Section~\ref{sec:conclusion} concludes.
