\section{Related Work}
\label{sec:related}

\paragraph{Vanilla speculative decoding.}
Leviathan \emph{et al.}~\cite{Leviathan2023} introduced the canonical
draft--verify scheme: a smaller \emph{draft} model proposes $k$ tokens
autoregressively and the larger \emph{target} verifies them in a single
parallel forward pass, with modified rejection sampling guaranteeing that the
output distribution is identical to that of the target. Chen
\emph{et al.}~\cite{Chen2023specsampling} concurrently proposed an equivalent
scheme. The expected per-cycle latency is
\begin{equation}
  L \;=\; \frac{T_{\text{draft}} + T_{\text{verify}}}{\tau},
  \qquad
  T_{\text{draft}} \;=\; k \cdot t_{\text{draft step}},
  \label{eq:specdec-latency}
\end{equation}
where $\tau \in [1, k{+}1]$ is the expected number of accepted tokens per
cycle. Wall-clock speedup $S = L_{\text{target}}/L$ therefore depends on the
draft/target cost ratio and the empirical token acceptance rate $\alpha$.
Stern \emph{et al.}~\cite{Stern2018blockwise} earlier explored blockwise
parallel decoding within a single model.

\paragraph{Reducing drafting cost.}
Medusa~\cite{Cai2024Medusa} eliminates the external draft by attaching
multiple prediction heads to the target model and verifying the resulting
candidate tree in one forward pass. The EAGLE
family~\cite{Li2024Eagle,Li2024Eagle2,Li2025Eagle3} keeps an external draft
but conditions it on the target's hidden features, reaching $\tau\approx 3$
and $S\approx 2\times$. Despite these refinements, $T_{\text{draft}}$
remains linear in $k$, which forces EAGLE-3 to a single transformer layer
and caps achievable speedups at $\sim$$2$--$3\times$ on strong GPUs.

\paragraph{Diffusion-based drafters.}
Recent work attacks the linear-in-$k$ drafting bottleneck directly.
DiffuSpec~\cite{Li2025DiffuSpec} and SpecDiff-2~\cite{Sandler2025SpecDiff2}
employ pre-trained diffusion LLMs as parallel drafters but rely on
$\sim$$7$\,B drafters whose memory footprint and per-step latency are
prohibitive in single-stream serving. PARD~\cite{An2025PARD} trains small
autoregressive models to mimic diffusion-style parallel generation but caps
at $\sim$$3\times$. DFlash~\cite{Chen2026dflash} closes this gap with a
lightweight (5-layer) \emph{block-diffusion} drafter conditioned on target
hidden features, reporting up to $5\times$ end-to-end speedup over
autoregressive baselines on H200/B200 hardware.

\paragraph{Position of this work.}
Our study is the first, to our knowledge, to combine two contributions in
the same protocol: (i) a controlled empirical evaluation of vanilla
speculative decoding on consumer-class hardware (RTX 5090,
Qwen2.5-7B-Instruct quantised to 8-bit via
\texttt{bitsandbytes}~\cite{Dettmers2022Int8}); and (ii) DriftDiffuse, a
small ($\sim$30\,M parameters) bidirectional masked-diffusion drafter that
operates within the consumer-hardware budget while breaking the
linear-in-$k$ drafting cost. We do not compete with DFlash's H200-class
numbers; we instead establish the operating regime in which a parallel
drafter remains useful when the AR draft is already well-tuned.
