\section{Experimental Protocol}
\label{sec:protocol}

\subsection{Datasets and sample sizes}

We evaluate on three benchmarks spanning reasoning, knowledge, and summarization:
\begin{enumerate}
    \item \textbf{GSM8K}~\cite{Cobbe2021} test subset: 300 samples.
    \item \textbf{MMLU}~\cite{Hendrycks2020} subset: 500 samples (5 subjects $\times$ 100 each).
    \item \textbf{CNN/DailyMail}~\cite{Narayan2018} subset: 200 samples.
\end{enumerate}

\noindent\textbf{Sampling policy:} Fixed random seed $= 42$, stratified sampling where applicable, and a frozen sample-ID manifest generated before experimentation. Each output CSV records seed and sample identifier, enabling paired analysis at sample level.

\subsection{Prompt set}

We use the following fixed prompt templates for each task:

\smallskip\noindent\textbf{GSM8K:}
\begin{quote}\small
You are a careful math assistant. Solve the problem step by step and end with
Final Answer: \textless number\textgreater.\\
Question: \{question\}
\end{quote}

\noindent\textbf{MMLU:}
\begin{quote}\small
You are a knowledgeable assistant. Choose exactly one option.\\
Question: \{question\}\\
Options: A.~\{A\} B.~\{B\} C.~\{C\} D.~\{D\}\\
Answer:
\end{quote}

\noindent\textbf{CNN/DailyMail:}
\begin{quote}\small
Summarize the following article in 3--4 sentences, preserving key facts and entities.\\
Article: \{article\}\\
Summary:
\end{quote}
\paragraph{Objective and hypotheses.}
The study quantifies practical speculative-decoding gains on consumer hardware
under fixed data and prompt controls.
We test four hypotheses:
\begin{itemize}
    \item \textbf{H1:} Same-family speculative decoding yields positive net speedup
        ($S>1$) over autoregressive decoding.
    \item \textbf{H2:} Increasing draft length from $k{=}4$ to $k{=}8$ improves
        speedup, while $k{=}16$ may reduce gains due to additional drafting
        and verification overhead.
    \item \textbf{H3:} A larger draft (1.5B) achieves higher acceptance and higher
        net speed than a smaller draft (0.5B).
    \item \textbf{H4:} Deterministic decoding is more runtime-efficient and quality
        stable than stochastic decoding for the same draft and $k$.
\end{itemize}

\paragraph{Hardware and software controls.}
All experiments run on one NVIDIA RTX~5090 Laptop (24GB VRAM) with a fixed software
stack (PyTorch, Transformers, and tokenizer settings pinned by the project
environment). We use the same runtime entry points and prompt templates for
all configurations, and execute runs serially to avoid cross-run contention.

\paragraph{Model pairing.}
Target model: Qwen2.5-3B-Instruct.
Draft models: Qwen2.5-0.5B-Instruct and Qwen2.5-1.5B-Instruct.
All models are same-family to minimise tokenizer mismatch and style drift.

\paragraph{Task suite and manifests.}
We evaluate on fixed manifests reused across all runs:
\begin{itemize}
    \item GSM8K arithmetic QA (300 samples),
    \item MMLU subset with 5 subjects (500 samples),
    \item CNN/DailyMail summarisation (200 samples).
\end{itemize}
Each configuration therefore processes 1{,}000 prompts.

\paragraph{Decoding regimes and grid.}
Two regimes are evaluated:
\begin{itemize}
    \item \textbf{Deterministic:} greedy decoding ($T{=}0.0$, top-$p{=}1.0$, \texttt{do\_sample=False}).
    \item \textbf{Stochastic:} sampling with $T{=}0.7$, top-$p{=}0.9$, no top-$k$ truncation, and \texttt{do\_sample=True}.
\end{itemize}
For each draft model, we sweep $k\in\{4,8,16\}$, producing
12 speculative configurations. We also run one autoregressive baseline per
regime for anchored latency and throughput reporting.

\paragraph{Endpoints.}
The primary endpoint is paired wall-clock speedup
$S = L_{\text{AR}}/L_{\text{spec}}$ as reported by the experiment summary
artifact for each draft--$k$--regime combination.
Secondary endpoints include acceptance rate $\alpha$, effective accepted block
size $B_{\text{eff}}$, mean per-sample latency, throughput (tokens/s), and
task-level quality deltas. Camera-ready statistical reporting uses paired
bootstrap confidence intervals (10{,}000 resamples) and one-sided tests for
$S{>}1$, with corrected pairwise winner tests.
