\section{Implementation and Reproducibility}
\label{sec:impl}

\paragraph{Code organisation.}
The experiment pipeline is implemented under the project source directory with
modular components for configuration, data loading, autoregressive baselines,
speculative decoding, metric aggregation, and runtime orchestration.
The notebook driver executes these modules in fixed phase order and writes
artifacts to the results directory.

\paragraph{Model loading and precision.}
In the current configuration, target and draft models are loaded in FP16 for
the main sweep, with quantisation controls retained in configuration for
alternative runs. Tokenizer and generation limits are held constant across
baselines and speculative runs to preserve paired comparability.

\paragraph{Baselines.}
We run one autoregressive baseline per regime (deterministic and stochastic)
with the same manifests and output-length constraints as speculative runs.
Each baseline artifact stores per-sample latency, output length, and task
predictions.

\paragraph{Speculative decoding implementation.}
The speculative loop follows standard token-level draft--verify logic.
At each cycle, the draft proposes up to $k$ tokens autoregressively; the
target verifies the proposal in one pass over the extended prefix; the accepted
prefix is committed up to the first mismatch; and one correction token is added
from the target distribution when needed. This procedure preserves target-level
generation semantics while exposing acceptance statistics needed for analysis.

\paragraph{Regimes and outputs.}
The same verification core is used for deterministic and stochastic modes,
with only sampling policy changed by regime parameters.
Each configuration writes a dedicated CSV artifact with per-sample runtime,
token counts, and quality-relevant outputs.

\paragraph{Aggregation and summary artifact.}
The metrics module merges paired records by sample identifier and computes
speedup $S$, acceptance $\alpha$, effective accepted block size
$B_{\text{eff}}$, latency, throughput, and task-level quality deltas.
The main reporting table used in this paper is the consolidated summary CSV
(\texttt{all\_configs\_summary.csv}).

\paragraph{Reproducibility controls and caveat.}
Runs are executed serially on one GPU, with fixed manifests and fixed decode
limits. We preserve raw per-configuration artifacts and derived summaries.
Because some baseline files were rerun after initial summary generation,
absolute baseline totals and summary-implied baselines may differ across
artifact sets; this paper therefore treats configuration ranking and paired
summary metrics as primary evidence, and reports this mismatch explicitly as a
threat to strict replay equivalence. For camera-ready reproducibility, reported
tables are tied to a coherent run family identified by manifest snapshot,
configuration state, and generation artifacts produced in a single execution
window. We release sufficient metadata and commands to regenerate all tables
from raw CSV outputs.
