\section{Discussion}
\label{sec:discussion}

\subsection{A systems study on consumer hardware}
The community-reported $2$--$5\times$ speedups for vanilla speculative
decoding are obtained on data-center GPUs with unquantised drafters.
Our consumer-hardware result of $S{=}1.65\times$ at $k{=}8$ is therefore
not an anomaly but a faithful read of the regime in which most
single-stream LLM serving actually happens: the
$T_{\text{verify}}$ term is inflated by \texttt{int8} dequantisation, while
$T_{\text{draft}}$ benefits proportionally less from the same kernel
fusions used in production stacks. Combined with the empirical observation
that $\bar{B}_{\text{eff}}$ tracks the geometric upper bound
$(1{-}\alpha^{k+1})/(1{-}\alpha)$ within $3\%$
(Section~\ref{sec:results}), this means that on RTX~5090-class hardware
the binding term is no longer acceptance: it is $T_{\text{draft}}=k\cdot
t_{\text{step}}$. This is the regime in which a parallel drafter such as
DriftDiffuse becomes structurally attractive
(Section~\ref{sec:driftdiffuse}).

\subsection{When is each draft size preferable?}
At $k{=}4$ the per-block overhead is not amortised and the 0.5B draft
under-performs the baseline ($S{=}0.77$). At $k{\geq}8$ the 0.5B draft
strictly dominates the 1.5B draft at every $k$, because the $\sim$$3\times$
parameter advantage of the 1.5B draft does not compensate for its $\sim$$3\times$
draft-step latency at the acceptance rates we measure
($\bar{\alpha}\!\in\![0.30,0.45]$). The practical recipe on this hardware
is therefore: use the smallest draft that retains $\bar{\alpha}\!\geq\!0.35$
and push $k$ to $8$ or $16$.

\subsection{Quality preservation}
Across all 12 speculative configurations, $|\Delta Q|\!\leq\!1.0$ point on
GSM8K and MMLU and $|\Delta\text{ROUGE-L}|\!\leq\!0.5$ on CNN/DailyMail.
This matches the theoretical guarantee of modified rejection sampling
under deterministic decoding and confirms it empirically under stochastic
decoding. The quality story is decoupled from the speedup story.

\section{Threats to validity}
\label{sec:threats}
\textbf{Model-family dependence.} Acceptance is conditional on the
draft and target sharing a tokenizer and pretraining distribution. Cross-family
pairings (e.g.\ Llama draft, Qwen target) typically yield lower $\alpha$ and
should not be assumed to inherit our $S$.
\textbf{Hardware dependence.} Absolute latencies will shift on H100/H200,
where lower-precision target kernels recover proportionally more of
$T_{\text{verify}}$ and reduce the relative gain of speculation.
\textbf{Prompt sensitivity.} GSM8K results are sensitive to chain-of-thought
phrasing; the prompt templates are pinned in \texttt{config.py}.
\textbf{Metric limits.} ROUGE-L on CNN/DailyMail does not capture
factuality, so $\Delta Q$ on summarisation is reported as a necessary but
not sufficient quality witness.

\section{Conclusion}
\label{sec:conclusion}

We presented a controlled empirical study of vanilla speculative decoding
on consumer hardware and used its findings to motivate a parallel
drafter. On a single RTX~5090 with an \texttt{int8} Qwen2.5-7B target,
the best autoregressive configuration delivers a $1.65\times$ wall-clock
speedup at $k{=}8$ with quality changes bounded within $1.0$ point. The
empirical effective block size matches the geometric upper bound within
$3\%$, isolating the linear-in-$k$ draft cost as the binding term. We
introduced \emph{DriftDiffuse}, a $\sim$30\,M-parameter masked-diffusion
drafter that proposes $k$ tokens in $n_{\text{denoise}}{\ll}k$ parallel
steps while remaining compatible with the standard verifier, and we
released the full pipeline together with frozen sample manifests and
environment pins. We see two natural extensions: training DriftDiffuse on
the target's own outputs at $k{\in}\{16,32\}$ to validate the predicted
break-even, and a cross-hardware study (RTX~5090, A100, H200) that
isolates how much of the consumer-hardware speedup ceiling is structural
rather than implementation-bound.
