\section{DriftDiffuse Extension (Prospective)}
\label{sec:driftdiffuse}

The current empirical study identifies a practical optimum at moderate block
size ($k{=}8$). Beyond this point, accepted block length continues to increase,
but speedup declines, indicating that additional drafting cost and verification
work are no longer offset by acceptance gains. This motivates a targeted
extension: replacing autoregressive drafting with a parallel drafter that can
propose a full block in fewer sequential steps.

\subsection{Design objective}

Let speculative-cycle latency be
\begin{equation}
  L = \frac{T_{\text{draft}} + T_{\text{verify}}}{\tau},
\end{equation}
where $\tau$ is expected accepted tokens per cycle. In autoregressive drafting,
$T_{\text{draft}}$ scales approximately linearly with $k$. A diffusion-style
parallel drafter aims to reduce this term by generating block tokens through a
small fixed number of denoising steps, making draft latency less sensitive to
$k$.

\subsection{Proposed approach}

We propose DriftDiffuse as a lightweight bidirectional masked-diffusion
drafter compatible with standard target-side verification. The intended
pipeline is:
\begin{enumerate}
  \item initialize a length-$k$ masked block conditioned on the accepted prefix,
  \item iteratively denoise in a small number of parallel refinement steps,
  \item submit the resulting draft block to the existing verifier,
  \item apply the same acceptance/rejection logic used in the current protocol.
\end{enumerate}

This design preserves verifier compatibility and allows direct comparison with
the autoregressive draft baseline under the same data manifests, prompts,
metrics, and hardware.

\subsection{Evaluation plan}

The extension will be evaluated as a strict incremental study rather than a new
protocol. We will retain:
\begin{itemize}
  \item the same target model and task manifests,
  \item the same deterministic and stochastic regimes,
  \item the same reporting metrics ($S$, $\alpha$, $B_{\text{eff}}$, quality deltas),
  \item the same artifact schema for per-configuration CSV outputs.
\end{itemize}

Primary comparison points are the current best autoregressive settings
($k{=}8$ deterministic) and high-$k$ settings ($k{=}16$) where the present
pipeline shows diminishing returns.

\subsection{Status}

At the time of writing, DriftDiffuse is a planned extension and is not included
in the quantitative results of Section~\ref{sec:results}. We include it here to
make the research trajectory explicit and to separate completed evidence from
forward-looking engineering work.
